\documentclass{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{geometry}

\geometry{margin=1in}

\title{Lecture Summary: IPL Powerplay Overs and Statistical Exploration}
\author{}
\date{}

\begin{document}

\maketitle

\section*{Source: Week 11 Lec 03.pdf}

\section*{Key Points}

\begin{itemize}
  \item \textbf{Focus on IPL Powerplay Overs:}
    \begin{itemize}
      \item The IPL powerplay covers the first six overs of each innings, characterized by field restrictions allowing only two players outside the inner circle.
      \item These overs are marked by high scoring potential and relatively consistent initial conditions across teams, making them suitable for statistical study.
    \end{itemize}

  \item \textbf{Data Collection:}
    \begin{itemize}
      \item Data from 1600 innings was extracted, capturing details like venue, innings, target, team, and runs scored in each delivery of the powerplay overs.
      \item Summarized metrics for each over include:
        \begin{itemize}
          \item Number of deliveries ($d_i$).
          \item Total runs scored ($r_i$).
          \item Wickets lost ($w_i$).
        \end{itemize}
    \end{itemize}

  \item \textbf{Statistical Observations:}
    \begin{itemize}
      \item \textbf{Average Runs by Over:}
        \[
          \text{1st Over: } 6, \quad \text{2nd Over: } 7, \quad \text{6th Over: } 8.3 \, (\text{highest}).
        \]
      \item \textbf{Standard Deviation of Runs by Over:}
        \[
          \text{Stabilizes at around 4-4.5 runs}.
        \]
      \item \textbf{Team-Specific Observations:}
        \begin{itemize}
          \item Delhi Daredevils (7.48 in 6th over) improved significantly as Delhi Capitals (8.94).
          \item Top performers: Sunrisers Hyderabad, Chennai Super Kings.
        \end{itemize}
      \item \textbf{Impact of Wickets in Over 5 on Runs in Over 6:}
        \begin{itemize}
          \item 0 wickets: $8.77$ average runs.
          \item 1 wicket: $6.7$ average runs.
          \item 2 wickets: Limited data, average drops to $7.6$.
        \end{itemize}
    \end{itemize}

  \item \textbf{Exploring the Data:}
    \begin{itemize}
      \item Tools like Google Sheets and Python are used to analyze and visualize patterns.
      \item Examples include:
        \begin{itemize}
          \item Histograms for runs scored distribution.
          \item Scatter plots, e.g., runs in over 6 vs. wickets in over 5.
        \end{itemize}
    \end{itemize}

  \item \textbf{Suggested Activity:}
    \begin{itemize}
      \item Students are encouraged to collect and analyze their own datasets.
      \item Avoid curated data sources like Kaggle; focus on raw, real-world data.
      \item Suggested size: At least 100 rows and 5 columns.
    \end{itemize}
\end{itemize}

\section*{Simplified Explanation}

\textbf{Powerplay Data:}
First six overs of IPL innings show high scoring potential, with field restrictions influencing play.

\textbf{Statistical Insights:}
- Runs increase progressively through overs, peaking in the 6th over.
- Team performance varies slightly, with top teams showing higher averages.
- Wickets lost in earlier overs significantly impact scoring potential in subsequent overs.

\textbf{Exploration:}
Students are encouraged to analyze data using tools like Python or spreadsheets, focusing on descriptive statistics and visualization.

\section*{Conclusion}

In this lecture, we:
\begin{itemize}
  \item Introduced IPL powerplay overs as a case study for statistical analysis.
  \item Highlighted patterns and insights derived from the data.
  \item Emphasized the importance of hands-on exploration with real-world datasets.
\end{itemize}

This example bridges theoretical statistics with practical applications in sports analytics.

\end{document}
